---
title: "The impact of NHI on health service utilization in Lao PDR"
author: "Sekeun Daniel Yu (yus109@mcmaster.ca)"
date: "April/2025"

---

```{r knitr, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)

```

# Setup
<br>
```{r setup, echo=TRUE, message=FALSE, results="hide"}

# Set working directory
rm(list = ls(all.names = TRUE))
#install.packages("rstudioapi")
library(rstudioapi)
setwd(dirname(rstudioapi::getSourceEditorContext()$path))
getwd()

# Install and load packages
source("lao_nhi_0_package.r", echo = TRUE)

```
<br>

# Load data
<br>
```{r load, echo=TRUE, message=FALSE, results="hide"}

data.cln <- read_excel("./data export/02_clean_8_cln_q_dl.xlsx") # district
warnings()
str(data.cln)

data <- data.cln

# Check number of district
data %>% 
  filter(faclevel != "PH") %>% 
  summarise(count = n_distinct(district)) # total district; 148

data %>% 
  filter(faclevel != "PH", treat == 1) %>% 
  summarise(count = n_distinct(district)) # total district with NHI; 139

data %>% 
  group_by(province) %>% 
  filter(faclevel != "PH") %>% 
  summarise(count = n_distinct(district))

data %>% 
  group_by(province, district) %>% 
  summarize(n = n()) %>% 
  print(n = Inf)

# Filter data
data <- data %>% 
  filter(
    !(province %in% c("15 Xekong")), # Drop NHI pilot province
    !(district %in% c("1008 Met", "1009 Viangkham", "1011 Mun",
                      "1306 Nong", "1408 Samouay")), # Drop NHI pilot districts
    !(district %in% c("0201 Phongsali")), # Drop district without CU5 in Q4/2015
    !str_detect(faclevel, "PH"), # Remove provincial hospitals
    !(quarter %in% c("2015.1", "2015.2", "2015.3")) # Quarters to drop
  ) 

# District with PH or HC only
# PH: XX00
# PSL 0205 Bounnua: HC only
# LNT 0301 Namtha: HC only
# ODX 0401 Xai: HC only
# HPN 0701 Xamnua: HC only
# XYB 0801 Xainyabouli: HC only
# XKG 0901 Pek: HC only
# BKX 1101 Pakxan: HC only
# KMN 1201 Thakhek: HC only
# SVK 1301 Kaysone Phomvihane: HC only
# SRV 1401 Salavan: HC only 
# SEK 1501 Lamam: HC only
# CPS 1601 Pakxe: HC only 
# ATP 1702 Samakhixai: HC only 
# XSB 1801 Anouvong: HC only 

# 1210 Khounkham - excluded in facility-level analysis but included in district-level analyses

# Excluding these 12 districts with HC only (due to the presence of PH), increases the DID estimates by about 0.05 points.
#data <- subset(data,District!="0205 Bounnua") # HC only
#data <- subset(data,District!="0301 Namtha") # HC only
#data <- subset(data,District!="0401 Xai") # HC only
#data <- subset(data,District!="0701 Xamnua") # HC only
#data <- subset(data,District!="0801 Xainyabouli") # HC only
#data <- subset(data,District!="0901 Pek") # HC only
#data <- subset(data,District!="1201 Thakhek") # HC only
#data <- subset(data,District!="1301 Kaysone Phomvihane") # HC only
#data <- subset(data,District!="1401 Salavan") # HC only
#data <- subset(data,District!="1601 Pakxe") # HC only
#data <- subset(data,District!="1702 Samakhixai") # HC only
#data <- subset(data,District!="1801 Anouvong") # HC only

# Create directories
if(!dir.exists("./output")) {dir.create("./output")}
if(!dir.exists("./graph")) {dir.create("./graph")}

```
<br>

# Table 1. Baseline characteristics
<br>
```{r tbl_1, echo=TRUE, warning=TRUE}

# package tableone
# http://www.danieldsjoberg.com/gtsummary-weill-cornell-presentation/#14
        
colnames(data)
str(data)

# District-level population characteristics
table1a <- data %>%
 filter(quarter == 2015.4) %>%                     # filter unique quarter
 filter(!group == 0) %>%                           # remove never treated
 select(group, facnum, imps, litr, semp, urba) %>%
 mutate(across(imps:urba, ~.x * 100)) %>%
 mutate(treat_g = ifelse(group == 9, "Control", "Intervention")) %>%
 tbl_strata(strata = treat_g,
            ~tbl_summary(.x,
                         by = group,
                         statistic = list(
                           facnum ~ "{sum}",
                           facnum ~ "{mean} ({sd})",
                           semp ~ "{mean} ({sd})",
                           urba ~ "{mean} ({sd})",
                           imps ~ "{mean} ({sd})",
                           litr ~ "{mean} ({sd})"),
                         label = list(
                           facnum ~ "Total facilities, No.",
                           facnum ~ "Mean facilities, No.",
                           semp ~ "Self-employed population, %",
                           urba ~ "Urban population, %",
                           imps ~ "Improved sanitation, % ",
                           litr ~ "Literacy among aged 15-64, %"),
                         digits = list(all_continuous() ~ 1)) %>%
              modify_header(all_stat_cols() ~ "**{level}**<br>(n={n})") %>%
              modify_footnote(update = everything() ~ NA))  # remove footnote 1
# modify_column_alignment(columns = everything(), align = "center")

table1a
show_header_names(table1a)

# imps: Percentage of population using improved sanitation
# litr: Literacy rate 15-64 old
# semp: Self-employment rate
# urba: Percentage of urban population

pnumdnum <- data %>% 
  group_by(group) %>%
  summarize(pcount = n_distinct(province),
            dcount = n_distinct(district)) %>%
  mutate(pcount = as.numeric(pcount),
         dcount = as.numeric(dcount))

pdcount <- data %>% 
  group_by(group) %>%
  summarize(pcount = n_distinct(province),
            dcount = n_distinct(district)) %>%
  mutate(pcount = as.numeric(pcount),
         dcount = as.numeric(dcount),
         p_by_d = pcount / dcount)

table1 <- data %>%
 filter(quarter == 2016.2) %>%  # filter unique quarter
 filter(!group == 0) %>%        # remove never treated
 mutate(facnum2 = facnum) %>%
 left_join(pdcount, by = c("group")) %>%
 select(p_by_d, district, group, facnum, facnum2, semp, urba, imps, litr) %>%
 mutate(district = 1) %>%
 mutate(across(semp:litr, ~.x * 100))
str(table1)

table1c <- table1 %>%
 filter(group %in% c(5, 6, 8, 9)) %>%  # remove never treated
 tbl_summary(by = group,
             statistic = list(
               p_by_d ~ "{sum}",
               district ~ "{sum}",
               facnum ~ "{sum}",
               facnum2 ~ "{mean} ({sd})",
               semp ~ "{mean} ({sd})",
               urba ~ "{mean} ({sd})",
               imps ~ "{mean} ({sd})",
               litr ~ "{mean} ({sd})"),
             label = list(
               p_by_d ~ "Province, No.",
               district ~ "District, No.",
               facnum ~ "Total facilities, No.",
               facnum2 ~ "Mean facilities, No.",
               semp ~ "Self-employed population, %",
               urba ~ "Urban population, %",
               imps ~ "Improved sanitation, % ",
               litr ~ "Literacy among aged 15-64, %"),
             digits = list(
               p_by_d ~ 0,
               district ~ 0,
               facnum ~ 0,
               facnum2 ~ 1,
               semp ~ 1,
               urba ~ 1,
               imps ~ 1,
               litr ~ 1),
             type = list(p_by_d ~ 'continuous',
                         district ~ 'continuous')) %>%
#                      all_continuous() ~ 1)) %>%
 modify_header(label = "**Variables**",
               stat_1 ~ "G5\n(q4/2016)\nn={n}",
               stat_2 ~ "G6\n(q1/2017)\nn={n}",
               stat_3 ~ "G8\n(q3/2017)\nn={n}",
               stat_4 ~ "G9\n(q4/2017)\nn={n}") %>%
 add_overall(last = TRUE, col_label = "All control\ngroups\nn={n}") %>%
 modify_footnote(update = everything() ~ NA)  # remove footnote 1
# modify_column_alignment(columns = everything(), align = "center")
table1c
show_header_names(table1c)  # province number not correct!

table1t <- table1 %>%
 filter(!group == 9) %>%  # remove never treated
 tbl_summary(by = group,
             statistic = list(
               p_by_d ~ "{sum}",
               district ~ "{sum}",
               facnum ~ "{sum}",
               facnum2 ~ "{mean} ({sd})",
               semp ~ "{mean} ({sd})",
               urba ~ "{mean} ({sd})",
               imps ~ "{mean} ({sd})",
               litr ~ "{mean} ({sd})"),
             label = list(
               p_by_d ~ "Province, No.",
               district ~ "District, No.",
               facnum ~ "Total facilities, No.",
               facnum2 ~ "Mean facilities, No.",
               semp ~ "Self-employed population, %",
               urba ~ "Urban population, %",
               imps ~ "Improved sanitation, % ",
               litr ~ "Literacy among aged 15-64, %"),
             digits = list(
               p_by_d ~ 0,
               district ~ 0,
               facnum ~ 0,
               facnum2 ~ 1,
               semp ~ 1,
               urba ~ 1,
               imps ~ 1,
               litr ~ 1),
             type = list(p_by_d ~ 'continuous',
                         district ~ 'continuous')) %>%
 modify_header(label = "**Variables**",
               stat_1 ~ "G4\n(q3/2016)\nn={n}",
               stat_2 ~ "G5\n(q4/2016)\nn={n}",
               stat_3 ~ "G6\n(q1/2017)\nn={n}",
               stat_4 ~ "G8\n(q3/2017)\nn={n}") %>%
 add_overall(last = TRUE, col_label = "All treated\ngroups\nn={n}") %>%
# modify_header(all_stat_cols() ~ "**{level}**\n(N={n})") %>%
 modify_footnote(update = everything() ~ NA)  # remove footnote 1
# modify_table_body(~.x %>% dplyr::relocate(stat_0, .after = stat_4))
# modify_column_alignment(columns = everything(), align = "center")
table1t
show_header_names(table1t)  # province number not correct!

tbl_merge <- 
  tbl_merge(
    list(table1c, table1t),
    tab_spanner = c("**Control**", "**Intervention**")) %>%
  modify_caption("Table 1. Characteristics of samples, by group")
tbl_merge

set_flextable_defaults(font.family = "Times New Roman")
tbl_merge %>%
 as_flex_table() %>%
 save_as_docx(path = "output/tbl_1_sample_characteristics.docx",
              pr_section = prop_section(
                page_size = page_size(orient = "landscape")))

# Add top header: Mean value (SD)
# Note: n is the total number of district-level units. Group (G) indicates group

```
<br>

## Figure 1. Population coverage by social health protection schemes from 2008 to 2021
<br>
```{r fig_1, echo=TRUE, message=FALSE, results="hide"}

data.shp <- read_excel("./data import/lao_shp_schemes.xlsx", 
                       sheet = "final") # district
data.shp <- data.shp %>% 
  pivot_longer(cols = -"scheme", names_to = "year", values_to = "population")

tpop <- data.shp %>%
  filter(scheme == "Total population") %>%
  select(-scheme) %>%
  rename(total_population = population, year = year)

data.shp <- data.shp %>%
  filter(scheme != "Total population") %>%
  left_join(tpop, by = "year") %>%
  mutate(pct = population / total_population)

data.shp %>% distinct(scheme)

scheme_list <- c("Formal sector Defense",
                 "Formal sector Civil servants",
                 "Formal sector Private employees",
                 "Informal sector Health Equity Fund for the poor (HEF)",
                 "Free mothers (FMAT)",
                 "Free children under 5 (FCU5)",
                 "Informal sector Community-based health insurance (CBHI)",
                 "Informal sector National Health Insurance (NHI)")
scheme_label <- c("Formal sector: Military and Police",
                  "Formal sector: Civil servants (LSSO-SASS)",
                  "Formal sector: Private enterprise workers (LSSO-SSO)", 
                  "Informal sector: Health Equity Fund for the poor (HEF)",
                  "Informal sector: Free maternity (FMAT)",
                  "Informal sector: Free children under five (FCU5)",
                  "Informal sector: Community-based health insurance (CBHI)",
                  "Informal sector: National Health Insurance (NHI)")

data.shp <- data.shp %>%
  mutate(scheme2 = factor(scheme, 
                          levels = scheme_list, 
                          labels = scheme_label))

write_xlsx(data.shp, "./output/fig_1_population_scheme.xlsx")

hiro <- met.brewer("Hiroshige") # Red -> Blue
mypalette <- hiro[c(1:5, 7:9)]
#hoku <- met.brewer("Hokusai3")[2]
#paqu <- met.brewer("Paquin")[7]
#paqu <- met.brewer("Paquin")[8]
vgh3 <- met.brewer("VanGogh3")[3]
mypalette[1] <- vgh3
tpopn <- 7337783

# plot -------------------------------------------------------------------------

plot <- ggplot(
  data.shp, aes(fill = scheme2, y = population, x = year)) + 
  geom_bar(position = position_stack(reverse = TRUE), 
           color = "grey25", stat = "identity", lwd = 0.2) +
  scale_y_continuous(limits = c(0, tpopn), expand = c(0, 0), 
                     labels = scales::comma) +
  scale_fill_manual(values = rev(mypalette), breaks = scheme_label) +
  ylab("Population") + 
  xlab("Year") +
  annotate("text", x = 7.6, y = 3000000, label = "33%", size = 3.3) +
  geom_segment(aes(x = 7.7, y = 2750000, xend = 8, yend = 2400000),
               arrow = arrow(length = unit(0.05, "in")),
               color = "grey30") +
  annotate("text", x = 9.6, y = 7250000, label = "93%", size = 3.3) +
  geom_segment(aes(x = 9.7, y = 7000000, xend = 10, yend = 6650000),
               arrow = arrow(length = unit(0.05, "in")),
               color = "grey30") +
  theme_light() +
  theme(
    text = element_text(family = "Source Sans Pro"),
    panel.border = element_blank(),  
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(color = "black"),
    plot.title = element_text(color = "black", vjust = 1.1, size = 10),
    plot.margin = margin(t = 0.2, r = 0.05, b = 0.1, l = 0.1, unit = "in"),
    axis.title.x = element_text(color = "black", size = 10, 
                                margin = margin(0.15, 0, 0, 0, "in")),
    axis.title.y = element_text(color = "black", vjust = 3, size = 10),
    axis.ticks = element_line(color = "black"),
    axis.text.y = element_text(color = "black", size = 10),
    axis.text.x = element_text(color = "black", size = 10, vjust = 0.5),
    legend.position = 'bottom',
    legend.title = element_blank(),
    legend.key.size = unit(0.2, "in"),
    legend.key.spacing.y = unit(0.025, "in"),
    legend.text = element_text(size = 10)) + 
  guides(fill = guide_legend(ncol = 1, reverse = TRUE, byrow = TRUE))

png("./graph/fig_1_population_scheme.png", 
    units = "in", width = 6.3, height = 6.5, res = 400)
print(plot)
dev.off()

```

## Supplementary Figure 1. Timeline of NHI expansion from 2015 to 2019
<br>
```{r sfig_1, echo=TRUE, message=FALSE, results="hide"}

data.cln <- read_excel("./data export/02_clean_8_cln_q_dl.xlsx") # district

data.cln %>%
 filter(faclevel != "PH") %>%
 summarise(count = n_distinct(district))

data.cln %>%
 filter(faclevel != "PH") %>%
 filter(pnum != 1) %>%
 summarise(count = n_distinct(district))

data.num <- data.cln %>%
 filter(pnum != 1) %>%
 filter(faclevel != "PH")

data.num <- data.num %>%
  select(quarter, pnum, treat, group) %>%
  mutate(quarter = as.factor(quarter),
         group = as.factor(group))
unique(data.num$group)
str(data.num)

prov_label <- c("Q3 2016 (Attapeu, Luang Namtha, Xaysomboun)",
                "Q4 2016 (Salavanh)",
                "Q1 2017 (Bolikhamxay, Houaphanh, Xieng Khouang, Oudomxay, Phongsaly)",
                "Q2 2017 (Sekong)",
                "Q3 2017 (Savannakhet, Champasak)",
                "Q4 2017 (Bokeo, Khammouane, Luang Prabang, Sayabouly, Vientiane)")

data.num <- data.num %>%
  mutate(
    group = factor(group, labels = c("Q3 2016", "Q4 2016", "Q1 2017", 
                                     "Q2 2017", "Q3 2017", "Q4 2017")),
    quarter = factor(quarter, labels = c("Q4.2015", "Q1.2016", "Q2.2016", 
                                         "Q3.2016", "Q4.2016", "Q1.2017",
                                         "Q2.2017", "Q3.2017", "Q4.2017", 
                                         "Q1.2018", "Q2.2018", "Q3.2018", 
                                         "Q4.2018", "Q1.2019")),
    group2 = factor(group, labels = prov_label))

naqrt <- c("Q1.2015", "Q2.2015", "Q3.2015", "Q4.2015", "Q1.2016", "Q2.2016")
total <- data.num %>%
  group_by(quarter) %>%
  summarise(count = sum(treat)) %>%
  mutate(count = replace(count, quarter %in% naqrt, NA))

mypalette <- met.brewer("Hokusai3", 6, "discrete") # Paquin / Hiroshige
names(mypalette) <- prov_label

# Legend with province names (aggregated blocked), bar outline
data.num2 <- data.num %>%
  group_by(group2, quarter) %>%
  summarise(count = sum(treat)) %>%
  setNames(c("intv_group", "quarter", "count"))

write_xlsx(data.num2, "./output/sfig_1_district_num.xlsx")

plot <- ggplot(
  data.num2, aes(fill = intv_group, y = count, x = quarter)) + 
  geom_bar(position = position_stack(reverse = TRUE), 
           color = "grey25", stat = "identity", lwd = 0.2) +
  geom_text(aes(x = quarter, y = count + 5, label = count, fill = NULL), 
            data = total, size = 3, fontface = "italic") +
  scale_y_continuous(limits = c(0, 155), expand = c(0, 0)) +
  scale_fill_manual(
    values = mypalette,
    breaks = prov_label) +
  labs(fill = "NHI Introduction (province)", 
       y = "Number of districts", 
       x = "Quarter") +
  theme_light() +
  theme(
    text = element_text(family = "Source Sans Pro"), # Font style
    panel.border = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(color = "black"),
    plot.title = element_text(color = "black", vjust = 1.1, size = 10),
    plot.margin = margin(t = 0.2, r = 0.05, b = 0.1, l = 0.1, unit = "in"),
    axis.title.x = element_text(color = "black", size = 10, 
                                margin = margin(t = 0.15, unit = "in")),
    axis.title.y = element_text(color = "black", vjust = 2, size = 10),
    axis.ticks = element_line(color = "black"),
    axis.text.y = element_text(color = "black", size = 10),
    axis.text.x = element_text(color = "black", size = 9.5, 
                               angle = 45, vjust = 0.75, hjust = 0.75),
    legend.title = element_text(color = "black", size = 10, 
                                margin = margin(b = 0.07, unit = "in")),
    legend.key.size = unit(0.2, "in"),
    legend.key.spacing.y = unit(0.025, "in"),
    legend.text = element_text(size = 10)) +
  guides(fill = guide_legend(byrow = TRUE))

png(paste("./graph/sfig_1_district_num.png", sep = ""), units = "in",
    width = 9.5, height = 4.5, res = 500)
print(plot)
dev.off()

```
<br>

## Supplementary Figure 2. Propensity score distribution by control group
<br>
```{r sfig_2, echo=TRUE, fig.height=2.5, fig.width=5}

# Reference
# https://cran.r-project.org/web/packages/cobalt/vignettes/cobalt.html#why-cobalt
# https://ngreifer.github.io/WeightIt/articles/WeightIt.html

# overlap 1: treated vs never-treated group
# ------------------------------------------------------------------------------

data.ovl <- data %>%
  filter(quarter == 2015.4) %>%              
  mutate(psgroup = ifelse(group > 0, 1, 0))

pscore.tr <- stats::glm(psgroup ~ logfacn + semp + urba + litr + imps,
                        family = "binomial", data = data.ovl)
ps.fit <- as.vector(pscore.tr$fitted.values)
ps.fit2 <- pmin(ps.fit, 1 - 1e-16)  # Adjust fitted values to avoid extremes

stat.ovl <- WeightIt::weightit(
  psgroup ~ logfacn + semp + urba + litr + imps,
  data = data.ovl,
  method = "glm",
  estimand = "ATT")

bal.tab(stat.ovl)
bal.plot(stat.ovl, var.name = "prop.score", 
         which = "both", type = "histogram", mirror = TRUE)
         # "both" shows unajusted and adjusted results
bal.plot(stat.ovl, var.name = "prop.score", 
         which = "both", mirror = FALSE) # "unadjusted" only shows adjusted results


# overlap 2: treated vs last treated cohort as control
# ------------------------------------------------------------------------------

data.ovl2 <- data %>%
  filter(group > 0) %>%                         # drop group = 0
  filter(quarter == 2015.4) %>%                 # filter unique quarter
  mutate(psgroup = ifelse(group == 9, 0, 1))    # last cohort as control
        
stat.ovl2 <- WeightIt::weightit(
  psgroup ~ logfacn + semp + urba + litr + imps,
  data = data.ovl2,
  method = "glm",
  estimand = "ATT")

bal.tab(stat.ovl2)
bal.plot(stat.ovl2, var.name = "prop.score", 
         which = "both", type = "histogram", mirror = TRUE)
bal.plot(stat.ovl2, var.name = "prop.score",
         which = "both", mirror = FALSE) # "unadjusted" only shows adjusted results


# overlap 3: treated vs last treated cohort + not-yet-treated as control
# ------------------------------------------------------------------------------

data.ovl2 <- data %>%
  filter(group > 0) %>%                         # drop group=0
  filter(quarter == 2015.4) %>%                 # filter unique quarter
  mutate(psgroup = ifelse(group == 9, 0, 1))    # last cohort as control

data.ovlnyt <- data.ovl2 %>%                    # extract
  filter(psgroup > 0) %>%                       # filter nyt group
  filter(group > 4) %>%                         # exclude the first treated
  mutate(psgroup = ifelse(psgroup == 1, 0, 0))  # nyt cohort as control

data.ovl3 <- rbind(data.ovl2, data.ovlnyt)      # combine dataset
        
stat.ovl3 <- WeightIt::weightit(
  psgroup ~ logfacn + semp + urba + litr + imps,
  data=data.ovl3,
  method = "glm",
  estimand = "ATT")

summary(stat.ovl3)
bal.tab(stat.ovl3)
bal.plot(stat.ovl3, var.name = "prop.score", 
         which = "both", type = "histogram", mirror = TRUE)
bal.plot(stat.ovl3, var.name = "prop.score",
         which = "both", mirror = FALSE) # "unadjusted" only shows adjusted results
         
love.plot(stat.ovl3, binary = "std", thresholds = c(m = .1))

# overlap 4: treated vs never-treated + all not-yet-treated group
# ------------------------------------------------------------------------------

data.ovlnvt <- data %>%
  filter(group == 0) %>%                   # keep never-treated group
  filter(quarter == 2015.4) %>%
  mutate(psgroup = 0)
data.ovl4 <- rbind(data.ovl3, data.ovlnvt) # combine dataset

stat.ovl4 <- WeightIt::weightit(
  psgroup ~ logfacn + semp + urba + litr + imps,
  data = data.ovl4,
  method = "glm",
  estimand = "ATT")

bal.tab(stat.ovl4)
bal.plot(stat.ovl4, var.name = "prop.score",
         which = "both", type = "histogram", mirror = TRUE)
bal.plot(stat.ovl4, var.name = "prop.score",
         which = "both", mirror = FALSE) # "unadjusted" only shows adjusted results


# Plots ------------------------------------------------------------------------

plot.ovl1 <- cbind.data.frame(stat.ovl$ps, stat.ovl$weights, 
                              stat.ovl$treat, "nvt")
plot.ovl2 <- cbind.data.frame(stat.ovl2$ps, stat.ovl2$weights, 
                              stat.ovl2$treat, "last")
plot.ovl3 <- cbind.data.frame(stat.ovl3$ps, stat.ovl3$weights, 
                              stat.ovl3$treat, "lastnyt")

plot.ovl.all <- lapply(
 list(plot.ovl1, plot.ovl2, plot.ovl3),
 setNames, c("ps", "weight", "treat", "control_type")) %>%
 do.call(rbind, .) %>%
 mutate(
   control_type = factor(control_type,
                  levels = c("lastnyt", "last", "nvt"),
                  labels = c("Later treated\n(not yet + last treated)",
                             "Last treated", "Never treated")),
   group = factor(treat,
                   levels = c(0, 1),
                   labels = c("Control", "Intervention")))

write_xlsx(plot.ovl.all, "./output/sfig_2_overlap.xlsx")

theme.ps <- function() {
 theme_bw() +
 theme(
   text = element_text(family = "Source Sans Pro"),
   panel.grid.major.x = element_blank(),
   panel.grid.major.y = element_blank(),
   panel.grid.minor = element_blank(),
   panel.background = element_blank(),
   plot.title = element_text(color = "black", hjust = 0.5, size = 8,
                             family = "Source Sans Pro"),
   axis.title.x = element_text(color = "black", vjust = -1.1, size = 7),
   axis.title.y = element_text(color = "black", vjust = 2, size = 7),
   axis.text.x = element_text(size = 6),
   axis.text.y = element_text(size = 6),
   axis.ticks.y = element_line(color = "black", linewidth = 0.25),
   axis.ticks.x = element_line(color = "black", linewidth = 0.25),
   strip.text = element_text(color = "black", size = 7),
   strip.text.x = element_text(margin = margin(0.04, 0, 0.04, 0, "in")),
   strip.background = element_blank(),
   legend.title = element_blank(),
   legend.text = element_text(size = 7),
   legend.key.size = unit(0.15, "in"),
   legend.key.height = unit(0.05, "in"),
   legend.position = c(0.15, -0.14),
   legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "in"),
   legend.key.width = unit(0.05, "in"))
}

# Plot with legend, scale fixed
plot.ovl.mirror.fix <- ggplot() +
 geom_histogram(data = subset(plot.ovl.all, treat == 0),
                aes(x = ps, y = after_stat(count / sum(count)),
                    fill = "Control"),
                bins = 30, col = "black", lwd = 0.2) +
 geom_histogram(data = subset(plot.ovl.all, treat == 1),
                aes(x = ps, y = -after_stat(count / sum(count)),
                    fill = "Intervention"),
                bins = 30, col = "black", lwd = 0.2) +
 facet_wrap(~control_type, ncol = 3) +
 scale_y_continuous(name = "Proportion") +
 scale_x_continuous(name = "Propensity score",
                    limits = c(-0.1, 1.1), seq(0, 1, by = 0.2)) +
 scale_fill_manual(
   values = c("Control" = "#F8766D", "Intervention" = "#9bd1a4"),
   guide = guide_legend(reverse = TRUE, nrow = 1)) +
 theme.ps()

plot.ovl.mirror.fix

png("./graph/sfig_2_overlap.png",
    units = "in", width = 5, height = 2.5, res = 400)
print(plot.ovl.mirror.fix)
dev.off()

```
<br>

## F. Health service utilization trends
<br>
```{r pta, echo=TRUE, warning=TRUE}

# By intervention group/cohort
outname <- c("opo5r", "opu5r", "ipo5r", "ipu5r", "iplos", "delr")

table1_cohort <- data %>%
 filter(quarter >= 2015.4 & quarter <= 2017.3) %>%
 subset(., province != "01 Vientiane Capital")  # Drop Vientiane Capital

head(table1_cohort)

group_list <- list()
t_total_list <- list()
c_total_list <- list()
total_list <- list()

for (i in 1:length(outname)) {
 outn <- rlang::sym(outname[i])
 group_list[[i]] <- list()
 t_total_list[[i]] <- list()
 c_total_list[[i]] <- list()
 total_list[[i]] <- list()

 for (j in c(4, 5, 6, 8)) {
   group_list[[i]][[j]] <- list()
   t_total_list[[i]][[j]] <- list()
   c_total_list[[i]][[j]] <- list()

   # Intervention group
   group_t <- table1_cohort %>%
     mutate(group.new = j) %>%
     mutate(event.time = time - group.new) %>%
     filter(group == j) %>%
     group_by(group, event.time) %>%
     summarise(util.mean = mean(!!outn, na.rm = TRUE),
               util.sd = sd(!!outn, na.rm = TRUE),
               util.num = n()) %>%
     mutate(util.se = util.sd / sqrt(util.num),
            util.lower.ci = util.mean - qt(1 - (0.05 / 2), util.num - 1) * util.se,
            util.upper.ci = util.mean + qt(1 - (0.05 / 2), util.num - 1) * util.se) %>%
     mutate(intv = "Intervention")

   # Control group
   group_c <- table1_cohort %>%
     mutate(group.new = j) %>%
     mutate(event.time = time - group.new) %>%
     filter(group > j) %>%
     filter(event < 0) %>%
     group_by(event.time) %>%
     summarise(util.mean = mean(!!outn, na.rm = TRUE),
               util.sd = sd(!!outn, na.rm = TRUE),
               util.num = n()) %>%
     mutate(util.se = util.sd / sqrt(util.num),
            util.lower.ci = util.mean - qt(1 - (0.05 / 2), util.num - 1) * util.se,
            util.upper.ci = util.mean + qt(1 - (0.05 / 2), util.num - 1) * util.se) %>%
     mutate(group = j) %>%
     mutate(intv = "Control")

   group_tc <- rbind(group_t, group_c)
   group_tc$outcome <- outname[i]
   group_tc$util.sd <- NULL
   group_list[[i]][[j]] <- group_tc

   # Total
   t_total <- table1_cohort %>%
     mutate(group.new = j) %>%
     mutate(event.time = time - group.new) %>%
     filter(group == j)

   t_total_list[[i]][[j]] <- t_total

   c_total <- table1_cohort %>%
     mutate(group.new = j) %>%
     mutate(event.time = time - group.new) %>%
     filter(group > j) %>%
     filter(event < 0)

   c_total_list[[i]][[j]] <- c_total
 }

 # Combined all possible control units
 t_total_unlist <- unlist(t_total_list[i], recursive = FALSE)
 t_total_bind <- do.call("rbind", t_total_unlist)

 c_total_unlist <- unlist(c_total_list[i], recursive = FALSE)
 c_total_bind <- do.call("rbind", c_total_unlist)

 t_total_comb <- t_total_bind %>%
   group_by(event.time) %>%
   summarise(util.mean = mean(!!outn, na.rm = TRUE),
             util.sd = sd(!!outn, na.rm = TRUE),
             util.num = n()) %>%
   mutate(util.se = util.sd / sqrt(util.num),
          util.lower.ci = util.mean - qt(1 - (0.05 / 2), util.num - 1) * util.se,
          util.upper.ci = util.mean + qt(1 - (0.05 / 2), util.num - 1) * util.se) %>%
   mutate(intv = "Intervention")

 c_total_comb <- c_total_bind %>%
   group_by(event.time) %>%
   summarise(util.mean = mean(!!outn, na.rm = TRUE),
             util.sd = sd(!!outn, na.rm = TRUE),
             util.num = n()) %>%
   mutate(util.se = util.sd / sqrt(util.num),
          util.lower.ci = util.mean - qt(1 - (0.05 / 2), util.num - 1) * util.se,
          util.upper.ci = util.mean + qt(1 - (0.05 / 2), util.num - 1) * util.se) %>%
   mutate(intv = "Control")

 tc_total <- rbind(t_total_comb, c_total_comb)
 tc_total$outcome <- outname[i]
 tc_total$util.sd <- NULL
 total_list[[i]] <- tc_total
}

# Control group trends appear similar across cohorts but differ subtly in early event.time periods. In the later post-NHI period (event time 3–4), the trends are identical because units being used are small and identical, mainly from groups 8 and 9 (last treated). Calculations are correct.

group_unlist <- unlist(group_list, recursive = FALSE)
total_bind <- do.call("rbind", total_list)
group_bind <- do.call("rbind", group_unlist)

# write_xlsx(total_bind,"./output/sfig_2_pta_qrt_total.xlsx")
# write_xlsx(group_bind,"./output/sfig_3_pta_qrt_group.xlsx")

# Total
# Averaging all units directly vs. averaging by group with weights should yield 
# identical results

CS_compute <- group_bind %>%
 group_by(event.time, intv, outcome) %>%
 mutate(weight = util.num / sum(util.num)) %>%
 mutate(avgdyn = sum(util.mean * weight)) %>%
 dplyr::select(c("event.time", "intv", "outcome", "avgdyn")) %>%
 unique() %>%
 arrange(outcome, intv, event.time)

CS_crosscheck <- left_join(CS_compute, total_bind,
                           by = c("event.time", "intv", "outcome")) %>%
  dplyr::select(contains(c("avgdyn", "util.mean"))) %>%
  print(n = Inf)

# The results are identical
# Average of all units from each cohort by event time is equivalent to
# 1) average by cohort, and 2) average with weighting by unit size (CS code)


```

# Supplementary Figure 3. Quarterly health service utilization trends, overall
<br>
```{r sfig_3, echo=TRUE, warning=TRUE}

var_list <- c("opo5r", "ipo5r", "opu5r", "ipu5r", "iplos", "delr")
var_label <- c("Outpatient visits per 1000, aged ≥ 5",
               "Inpatient admissions per 100 000, aged ≥ 5",
               "Outpatient visits per 1000, aged < 5",
               "Inpatient admissions per 100 000, aged < 5",
               "Inpatient average length of stay",
               "Institutional birth rate")

total_bind <- total_bind %>%
  mutate(outcome2 = factor(outcome, 
                           levels = var_list, 
                           labels = var_label))

xl <- c(min(total_bind$event.time), # x-axis length / index
          max(total_bind$event.time))

write_xlsx(total_bind,"./output/sfig_3_pta_overall.xlsx")

plot <- ggplot(
  total_bind, aes(x = event.time, y = util.mean, 
                  group = intv, color = intv)) + 
  facet_wrap(~outcome2, ncol = 2, scales = "free") + 
  geom_vline(
    aes(xintercept = 0), linetype = "dashed", 
    color = "black", alpha = 0.9, linewidth = 0.05) +
  geom_ribbon(
    data = subset(total_bind, intv == "Intervention"),
    aes(ymin = util.lower.ci, ymax = util.upper.ci),
    fill = "#c5dec9", color = "#bad9bf", alpha = 0.65, 
    linetype = 1, linewidth = 0.025) +
  geom_ribbon(
    data = subset(total_bind, intv == "Control"),
    aes(ymin = util.lower.ci, ymax = util.upper.ci),
    fill = "#f5cdc6", color = "#f5beb5", alpha = 0.8, 
    linetype = 1, linewidth = 0.025) +
  geom_point(data = subset(total_bind, intv == "Control"), size = 0.6) +
  geom_line(data = subset(total_bind, intv == "Control"), linewidth = 0.3) +
  geom_point(data = subset(total_bind, intv == "Intervention"), size = 0.6) +
  geom_line(data = subset(total_bind, intv == "Intervention"), linewidth = 0.3) +
  scale_y_continuous(name = "Health service utilization") +
  scale_x_continuous(name = "Time since NHI introduction, quarter", 
                     breaks = xl[1]:xl[2]) +
  scale_color_manual(values = c("Control" = "#e85238", "Intervention" = "#006918"),
                     guide = guide_legend(reverse = TRUE, nrow = 1)) +
  theme_bw() +
  theme(
    text = element_text(family = "Source Sans Pro"),
    panel.border = element_rect(color = "black", linewidth = 0.4),
    panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
    plot.title = element_text(color = "black", size = 8),
    plot.margin = margin(t = 0.2, r = 0.2, b = 0.1, l = 0.1, unit = "in"),
    axis.title.x = element_text(color = "black", size = 7, 
                                margin = margin(0.15, 0, 0, 0, "in")),
    axis.title.y = element_text(color = "black", size = 7),
    axis.ticks = element_line(color = "black", linewidth = 0.2),
    axis.text = element_text(color = "black", size = 7),
    strip.text = element_text(color = "black", size = 7),
    strip.text.x = element_text(margin = margin(0.15, 0, 0.05, 0, "in")),
    strip.background = element_blank(),
    legend.title = element_blank(),
    legend.text = element_text(size = 6),
    legend.key.size = unit(0.2, "in"),
    legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "in"),
    legend.position = "inside", 
    legend.position.inside =  c(0.1, -0.067)
  )

png("graph/sfig_3_pta_overall.png", 
    units = "in", width = 5.1, height = 6, res = 500)
print(plot)
dev.off()

```

# Supplementary Figure 4. Quarterly health service utilization trends by intervention group
<br>
```{r sfig_4, echo=TRUE, warning=TRUE}

var_list <- c("opo5r", "ipo5r", "opu5r", "ipu5r", "iplos", "delr")
var_label_l <- c("Outpatient\nvisits\nper 1000,\naged ≥ 5",
                 "Inpatient\nadmissions\nper 100 000,\naged ≥ 5",
                 "Outpatient\nvisits\nper 1000,\naged < 5",
                 "Inpatient\nadmissions\nper 100 000, \naged < 5",
                 "Inpatient\naverage\nlength\nof stay",
                 "Institutional\nbirth\nrate")

group_bind <- group_bind %>%
  mutate(cohort = factor(group, 
                         levels = c(4, 5, 6, 8),
                         labels = c("NHI introduced in \nQ3/2016\nN=15",
                                    "NHI introduced in \nQ4/2016\nN=7",
                                    "NHI introduced in \nQ1/2017\nN=37",
                                    "NHI introduced in \nQ3/2017\nN=24")),
         outcome2 = factor(outcome, levels = var_list, labels = var_label_l))

xl <- c(min(group_bind$event.time), # x-axis length / index
          max(group_bind$event.time))

write_xlsx(group_bind,"./output/sfig_4_pta_cohort.xlsx")

plot <- ggplot(
  group_bind, aes(x = event.time, y = util.mean, group = intv, color = intv)) +
  facet_grid(outcome2 ~ cohort, scales = "free", switch = "y") +
  geom_vline(aes(xintercept = 0), linetype = "dashed", 
             color = "black", alpha = 0.9, linewidth = 0.05) +
  geom_ribbon(data = subset(group_bind, intv == "Intervention"),
              aes(x = event.time, ymin = util.lower.ci, ymax = util.upper.ci),
              fill = "#c5dec9", color = "#bad9bf", alpha = 0.65, 
              linetype = 1, linewidth = 0.025) +
  geom_ribbon(data = subset(group_bind, intv == "Control"),
              aes(x = event.time, ymin = util.lower.ci, ymax = util.upper.ci),
              fill = "#f5cdc6", color = "#f5beb5", alpha = 0.8, 
              linetype = 1, linewidth = 0.025) +
  geom_point(data = subset(group_bind, intv == "Control"), 
             aes(x = event.time, y = util.mean), size = 0.6) +
  geom_line(data = subset(group_bind, intv == "Control"), 
            aes(x = event.time, y = util.mean), linewidth = 0.3) +
  geom_point(data = subset(group_bind, intv == "Intervention"), 
             aes(x = event.time, y = util.mean), size = 0.6) +
  geom_line(data = subset(group_bind, intv == "Intervention"), 
            aes(x = event.time, y = util.mean), linewidth = 0.3) +
  scale_y_continuous(name = "Health service utilization") +
  scale_x_continuous(name = "Time since NHI introduction, quarter",
                     breaks = xl[1]:xl[2]) +
  scale_color_manual(values = c("Control" = "#e85238", 
                                "Intervention" = "#006918"),
                     guide = guide_legend(reverse = TRUE, nrow = 1)) +
  theme_bw() +
  theme(
    text = element_text(family = "Source Sans Pro"),
    panel.border = element_rect(color = "black", linewidth = 0.4),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    plot.title = element_text(color = "black", size = 8),
    plot.margin = margin(t = 0.1, r = 0.1, b = 0.1, l = 0, unit = "in"),
    axis.title.x = element_text(
      color = "black", size = 7, margin = margin(0.15, 0, 0, 0, "in")),
    axis.title.y = element_blank(),
    axis.ticks.y = element_line(color = "black", linewidth = 0.2),
    axis.ticks.x = element_line(color = "black", linewidth = 0.2),
    axis.text.y = element_text(color = "black", size = 7),
    axis.text.x = element_text(color = "black", size = 7),
    strip.placement = "outside",
    strip.text = element_text(color = "black", size = 7),
    strip.text.y.left = element_text(
      angle = 0, hjust = 0, 
      margin = margin(t = 0, r = 0, b = 0, l = 0.05, "in")),
    strip.text.x = element_text(
      margin = margin(0.15, 0, 0.05, 0, "in")),
    strip.background = element_blank(),
    legend.title = element_blank(),
    legend.text = element_text(size = 6),
    legend.key.size = unit(0.2, "in"),
    legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "in"),
    legend.position = "inside", 
    legend.position.inside =  c(0.05, -0.057)
  )

png("graph/sfig_4_pta_cohort.png",
    units = "in",width = 4.75, height = 7, res = 500)
print(plot)
dev.off()

```

