---
title: "The impact of NHI on health service utilization in Lao PDR"
author: "Sekeun Daniel Yu (yus109@mcmaster.ca)"
date: "April/2025"

---

```{r knitr, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)

```

# Setup
<br>
```{r setup, echo=TRUE, message=FALSE, results="hide"}

# Set working directory
rm(list = ls(all.names = TRUE))
#install.packages("rstudioapi")
library(rstudioapi)
setwd(dirname(rstudioapi::getSourceEditorContext()$path))
getwd()

# Install and load packages
source("lao_nhi_0_package.r", echo = TRUE)

# Create directories
if(!dir.exists("./graph")) {dir.create("./graph")}


var_list <- c("opo5r", "ipo5r", "opu5r", "ipu5r", "iplos", "delr")

# Label top
var_label_t <- c("Outpatient visits per 1000, aged ≥ 5",
               "Inpatient admissions per 100 000, aged ≥ 5",
               "Outpatient visits per 1000, aged < 5",
               "Inpatient admissions per 100 000, aged < 5",
               "Inpatient average length of stay",
               "Institutional birth rate")

# Label left
var_label_l <- c("Outpatient\nvisits\nper 1000,\naged ≥ 5",
                 "Inpatient\nadmissions\nper 100 000,\naged ≥ 5",
                 "Outpatient\nvisits\nper 1000,\naged < 5",
                 "Inpatient\nadmissions\nper 100 000, \naged < 5",
                 "Inpatient\naverage\nlength\nof stay",
                 "Institutional\nbirth\nrate")

```
<br>

# Supplementary Figure Layer and Theme (r6 x c4)
<br>
```{r sfig_theme, eval=FALSE, include=FALSE}

layer.64 <-
  list(
    geom_vline(aes(xintercept = 0), 
             linetype = "dashed", color = "black", 
             alpha = 0.9, linewidth = 0.05),
    geom_rect(aes(xmin = 0, xmax = Inf, 
                ymin = overall.conf.low, ymax = overall.conf.high), 
            fill = "#c5dec9", alpha = 0.15),
    geom_hline(aes(yintercept = 0), color = "black", linewidth = 0.175),
    geom_segment(aes(x = 0, xend = Inf, 
                   y = overall.conf.low, yend = overall.conf.low), 
               color = "#bad9bf", linewidth = 0.05, alpha = 0.5),
    geom_segment(aes(x = 0, xend = Inf, 
                   y = overall.conf.high, yend = overall.conf.high), 
               color = "#bad9bf", linewidth = 0.05, alpha = 0.5),
    geom_point(size = 0.6, color = "#006918"),
    geom_linerange(aes(x = event.time, ymin = conf.low, ymax = conf.high), 
                 size = 0.3, color = "#006918"),
    geom_segment(aes(x = 0, xend = Inf, 
                   y = overall.att, yend = overall.att), 
               color = "#d62b47", linewidth = 0.2, alpha = 0.5),
    geom_text(x = Inf, y = Inf, label = expression(Delta),
            family = "Source Sans Pro", color = "black",
            vjust = 1.5, hjust = 6, size = 2),
    geom_text(aes(x = x.pctchange, y = Inf, label = pctchange),
            family = "Source Sans Pro", color = "black",
            vjust = 1.5, hjust = 0.5, size = 2))

theme.64 <- function() {
  theme_bw() +
    theme(
      text = element_text(family = "Source Sans Pro"),
      panel.border = element_rect(color = "black", linewidth = 0.4),
      panel.grid.major.x = element_blank(),
      panel.grid.major.y = element_blank(),
      panel.grid.minor = element_blank(),
      panel.background = element_blank(),
      plot.title = element_text(color = "black", size = 8),
      plot.margin = margin(t = 0.1, r = 0.1, b = 0.1, l = 0, unit = "in"),
      axis.title.x = element_text(color = "black", size = 7, 
                                  margin = margin(0.15, 0, 0, 0, "in")),
      axis.title.y = element_blank(),
      axis.ticks.y = element_line(color = "black", linewidth = 0.2),
      axis.ticks.x = element_line(color = "black", linewidth = 0.2),
      axis.text.y = element_text(color = "black", size = 7),
      axis.text.x = element_text(color = "black", size = 7),
      strip.placement = "outside",
      strip.text = element_text(color = "black", size = 7),
      strip.text.y.left = element_text(angle = 0, hjust = 0, 
                                       margin = margin(t = 0, r = 0, b = 0, 
                                                       l = 0.05, "in")),
      strip.text.x = element_text(margin = margin(0.15, 0, 0.05, 0, "in")),
      strip.background = element_blank()
      # legend.title = element_blank(),
      # legend.text = element_text(size = 7),
      # legend.position = c(0.75, 0.1)
    )
}

```

# Main Figure 3. DiD analysis
<br>
```{r fig_3_main, eval=FALSE, include=FALSE}

# Load data
plot.dyn <- var_list %>%
  map(~ read_excel(paste0("./output/q_disl/", .x, "_aggte_dyn.xlsx"))) %>%
  bind_rows() %>%
  mutate(outcome2 = factor(outcome, 
                           levels = var_list, 
                           labels = var_label_t),
         across(c(overall.att, overall.se, 
                  overall.conf.low, overall.conf.high), 
                ~ ifelse(event.time < 0, NA, .)),
         pctchange = scales::percent(overall.change, accuracy = 0.1),
         pctchange = if_else(event.time != 0, NA, pctchange)) %>%
  filter(name == "uc.cls")

write_xlsx(plot.dyn, paste0("./output/", "fig_03_did_main.xlsx"))

# Plot -------------------------------------------------------------------------

xl <- range(plot.dyn$event.time) # x-axis length / index

plot <- 
  ggplot(plot.dyn, aes(x = event.time, y = estimate)) +
  facet_wrap(~outcome2, ncol = 2, scales = "free") +
  geom_vline(
    aes(xintercept = 0), linetype = "dashed", col = "black", 
    alpha = 0.9, linewidth = 0.05) +
  geom_rect(aes(xmin = 0, xmax = Inf, ymin = overall.conf.low, 
                ymax = overall.conf.high), fill = "#c5dec9", alpha = 0.15) +
  geom_hline(aes(yintercept = 0), col = "black", linewidth = 0.175) +
  geom_segment(aes(x = 0, xend = Inf, y = overall.conf.low, 
                   yend = overall.conf.low), color = "#bad9bf", 
               linetype = 1, linewidth = 0.05, alpha = 0.5) +
  geom_segment(aes(x = 0, xend = Inf, y = overall.conf.high, 
                   yend = overall.conf.high), color = "#bad9bf", 
               linetype = 1, linewidth = 0.05, alpha = 0.5) +
  geom_point(size = 0.6, color = "#006918") +
  geom_linerange(aes(x = event.time, ymin = conf.low, ymax = conf.high),
                 size = 0.3, color = "#006918") +
  geom_segment(aes(x = 0, xend = Inf, y = overall.att, yend = overall.att),
               color = "#d62b47", linewidth = 0.2, alpha = 0.5) +
  geom_text(x = 1.7, y = Inf, label = expression(Delta), 
            family = "Source Sans Pro", color = "black", 
            vjust = 1.5, size = 2) +
  geom_text(aes(x = 2.5, y = Inf, label = pctchange), 
            family = "Source Sans Pro", color = "black", 
            vjust = 1.5, size = 2) +
  scale_y_continuous(name = "Difference-in-differences estimates") +
  scale_x_continuous(name = "Time since NHI introduction, quarter", 
                     breaks = xl[1]:xl[2]) +
  theme_bw() +
  theme(
    text = element_text(family = "Source Sans Pro"),
    panel.border = element_rect(color = "black", linewidth = 0.4),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    plot.title = element_text(color = "black", size = 8),
    plot.margin = margin(t = 0.2, r = 0.2, b = 0.1, l = 0.1, unit = "in"),
    axis.title.x = element_text(color = "black", size = 7,
                                margin = margin(0.15, 0, 0, 0, "in")),
    axis.title.y = element_text(color = "black", size = 7),
    axis.ticks.y = element_line(color = "black", linewidth = 0.2),
    axis.ticks.x = element_line(color = "black", linewidth = 0.2),
    axis.text.y = element_text(color = "black", size = 7),
    axis.text.x = element_text(color = "black", size = 7),
    strip.text = element_text(color = "black", size = 7),
    strip.text.x = element_text(margin = margin(0.15, 0, 0.05, 0, "in")),
    strip.background = element_blank(),
    legend.title = element_blank(),
    legend.text = element_text(size = 7),
    legend.position.inside = c(0.75, 0.1))

png(paste0("./graph/", "fig_03_did_main.png"), 
    units = "in", width = 5.1, height = 6, res = 500)
print(plot)
dev.off()

```
<br>

# Supplementary Figure 5. DiD analysis by facility type
<br>
```{r sfig_5_facility, eval=FALSE, include=FALSE}

# Load data
read_data <- function(fl, vl) {
  map_dfr(vl, ~ {
    filepath <- paste0("./output/q_facl/", fl, 
                       "/", .x, "_aggte_dyn_", fl, ".xlsx")
    read_excel(filepath) %>%
      mutate(facl = fl)
  })
}

plot.fac <- map_dfr(c("dh", "hc"), 
                    ~ read_data(.x, var_list)) %>%
  mutate(outcome2 = factor(outcome,
                           levels = var_list,
                           labels = var_label_l),
         across(c(overall.att, overall.se, 
                  overall.conf.low, overall.conf.high),
                ~ ifelse(event.time < 0, NA, .)),
         pctchange = scales::percent(overall.change, accuracy = 0.1),
         pctchange = if_else(event.time != 0, NA, pctchange),
         payment = case_when(
           facl == "dh" & outcome %in% c("opo5r", "opu5r") ~ "FFS to Capitation",
           facl == "dh" & outcome %in% c("ipo5r", "ipu5r", "iplos", "delr") ~ 
             "FFS to Case-based payment",
           facl == "hc" ~ "FFS to Capitation",
           TRUE ~ NA),
         payment = if_else(is.na(pctchange), NA, payment),
         facl = factor(facl,
                       levels = c("dh", "hc"),
                       labels = c("District hospitals", "Health centers"))) %>%
  filter(name == "uc.cls")

write_xlsx(plot.fac, paste0("./output/", "sfig_05_did_facility.xlsx"))

# Plot -------------------------------------------------------------------------

xl <- range(plot.fac$event.time) # x-axis length / index

plot <- ggplot(
  plot.fac, aes(x = event.time, y = estimate)) + 
  facet_grid(outcome2 ~ facl, scales = "free", switch = "y") + 
  geom_vline(aes(xintercept = 0), linetype = "dashed", 
             color = "black", alpha = 0.9, linewidth = 0.05) +
  geom_hline(aes(yintercept = 0), color = "black", linewidth = 0.175) +
  geom_rect(aes(xmin = 0, xmax = Inf, ymin = overall.conf.low, 
                ymax = overall.conf.high), fill = "#c5dec9", alpha = 0.15) +
  geom_segment(aes(x = 0, xend = Inf, y = overall.conf.low, 
                   yend = overall.conf.low), color = "#bad9bf", 
               linewidth = 0.05, alpha = 0.5) +
  geom_segment(aes(x = 0, xend = Inf, y = overall.conf.high, 
                   yend = overall.conf.high), color = "#bad9bf", 
               linewidth = 0.05, alpha = 0.5) +
  geom_segment(aes(x = 0, xend = Inf, y = overall.att, yend = overall.att), 
               color = "#d62b47", linewidth = 0.2, alpha = 0.5) +
  geom_point(size = 0.6, color = "#006918") +
  geom_linerange(aes(ymin = conf.low, ymax = conf.high), 
                 size = 0.3, color = "#006918") +
  geom_text(x = 1.5, y = Inf, label = expression(Delta), 
            family = "Source Sans Pro", color = "black", vjust = 1.5, size = 2) +
  geom_text(aes(x = 2.7, y = Inf, label = pctchange), 
            family = "Source Sans Pro", color = "black", vjust = 1.5, size = 2) +
#  geom_text(aes(x = -7, y = Inf, label = payment), 
#            family = "Source Sans Pro", color = "black", vjust = 1.5, hjust = 0,
#            size = 1.8) +

  scale_x_continuous(name = "Time since NHI introduction, quarter", 
                     breaks = xl[1]:xl[2]) +
  theme_bw() +
  theme(
    text = element_text(family = "Source Sans Pro"),
    panel.border = element_rect(color = "black", linewidth = 0.4),
    panel.grid = element_blank(),
    plot.margin = margin(t = 0.1, r = 0.1, b = 0.05, l = 0, unit = "in"),
    axis.title.x = element_text(color = "black", size = 6, 
                                margin = margin(0.15, 0, 0, 0, "in")),
    axis.title.y = element_blank(),
    axis.ticks = element_line(color = "black", linewidth = 0.2),
    axis.text = element_text(color = "black", size = 6),
    strip.placement = "outside",
    strip.text = element_text(color = "black", size = 6),
    strip.text.y.left = element_text(
      angle = 0, hjust = 0,
      margin = margin(t = 0, r = 0, b = 0, l = 0.05, "in")),
    strip.text.x = element_text(margin = margin(0.15, 0, 0.05, 0, "in")),
    strip.background = element_blank(),
    legend.title = element_blank(),
    legend.text = element_text(size = 6),
    legend.position.inside = c(0.75, 0.1))

png(paste0("./graph/", "sfig_05_did_facility.png"), 
    units = "in", width = 4, height = 5.5, res = 500)
print(plot)
dev.off()

```
<br>

# Supplementary Figure 6. DiD analysis by district-level poverty status
<br>
```{r sfig_6_poverty, eval=FALSE, include=FALSE}

plot.pov <- read_excel("./output/sfig_06_did_poverty.xlsx")
plot.pov <- plot.pov %>%
  filter(pgroup != "Overall") %>%
  mutate(outcome2 = factor(outcome, 
                           levels = var_list, 
                           labels = var_label_t),
         across(c(oval.est, oval.se, 
                  oval.conf.low, oval.conf.high),
                ~ ifelse(event.time < 0, NA, .)),
         pctchange = scales::percent(oval.change, accuracy = 0.1),
         pctchange = if_else(event.time != 0, NA, pctchange))

write_xlsx(plot.pov, paste0("./output/", "sfig_06_did_poverty.xlsx"))

# Plot -------------------------------------------------------------------------

xl <- range(plot.pov$event.time) # x-axis length / index

plot <- ggplot(
  plot.pov, aes(x = event.time, y = estimate, color = pgroup, shape = pgroup)) + 
  facet_wrap(~outcome2, ncol = 2, scales = "free") +
  geom_vline(aes(xintercept = 0), linetype = "dashed", 
             color = "black", alpha = 0.9, linewidth = 0.05) +
  geom_hline(aes(yintercept = 0), color = "black", linewidth = 0.175) +
  geom_point(position = position_dodge2(width = 0.6), size = 0.9) +
  geom_linerange(aes(x = event.time, ymin = conf.low, ymax = conf.high), 
                 position = position_dodge2(width = 0.6), size = 0.3) +
  geom_text(x = 0.7, y = Inf, label = expression(Delta), 
            family = "Source Sans Pro", color = "#FC4E07", 
            vjust = 1.5, size = 2) +
  geom_text(data = subset(plot.pov, pgroup == "High poverty"),
            aes(x = 1.5, y = Inf, label = pctchange), 
            family = "Source Sans Pro", color = "#FC4E07", 
            vjust = 1.5, size = 2) +
  geom_text(x = 2.7, y = Inf, label = expression(Delta), 
            family = "Source Sans Pro", color = "#00AFBB", 
            vjust = 1.5, size = 2) +
  geom_text(data = subset(plot.pov, pgroup == "Low poverty"),
            aes(x = 3.5, y = Inf, label = pctchange), 
            family = "Source Sans Pro", color = "#00AFBB", 
            vjust = 1.5, size = 2) +
  scale_y_continuous(name = "Difference-in-differences estimates") +
  scale_x_continuous(name = "Time since NHI introduction, quarter",
                     breaks = xl[1]:xl[2]) +
  scale_color_manual(values = c("#FC4E07", "#00AFBB"),
                     guide = guide_legend(nrow = 1)) +
  theme_bw() +
  theme(
    text = element_text(family = "Source Sans Pro"),
    panel.border = element_rect(color = "black", linewidth = 0.4),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    plot.title = element_text(color = "black", size = 8),
    plot.margin = margin(t = 0.2, r = 0.2, b = 0.1, l = 0.1, unit = "in"),
    axis.title.x = element_text(color = "black", size = 7,
                                margin = margin(t = 0.15, unit = "in")),
    axis.title.y = element_text(color = "black", size = 7),
    axis.ticks = element_line(color = "black", linewidth = 0.2),
    axis.text = element_text(color = "black", size = 7),
    strip.text = element_text(color = "black", size = 7),
    strip.text.x = element_text(margin = margin(t = 0.15, b = 0.05, unit = "in")),
    strip.background = element_blank(),
    legend.title = element_blank(),
    legend.text = element_text(size = 6),
    legend.key.size = unit(0.2, "in"),
    legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "in"),
    legend.position = c(0.1, -0.065))

png(paste0("./graph/", "sfig_06_did_poverty.png"), 
    units = "in", width = 5.1, height = 6, res = 500)
print(plot)
dev.off()

```
<br>

# Supplementary Figure 7. DiD analysis by intervention group
<br>
```{r sfig_7_group, eval=FALSE, include=FALSE}

# Load data
plot.grp.all <- var_list %>%
  map(~ read_excel(paste0("./output/q_disl/", .x, "_attgt.xlsx"))) %>%
  bind_rows() %>%
  dplyr::select(-c("weight", "n.group"))

plot.grp <- var_list %>%
  map(~ read_excel(paste0("./output/q_disl/", .x, "_aggte_grp.xlsx"))) %>%
  bind_rows() %>%
  dplyr::select(-weight, -n.group) %>%
  filter(group != "Average") %>%
  mutate(group = as.numeric(group)) %>%
  select(group, name, pretrendtype, outcome, 
         estimate, std.error, conf.low, conf.high, change) %>%
  rename(att = estimate, se = std.error) %>%
  rename_at(vars(att:change), ~ paste0("overall.", .)) %>%
  mutate(pctchange = scales::percent(overall.change, accuracy = 0.1))

plot.grp <- plot.grp %>% 
  left_join(plot.grp.all,
            by = c("group","name","pretrendtype","outcome")) %>%
  mutate(event.time = time - group,
         group = factor(group, 
                        levels = c(4, 5, 6, 8),
                        labels = c("NHI introduced in \nQ3/2016\nN=15",
                                   "NHI introduced in \nQ4/2016\nN=7",
                                   "NHI introduced in \nQ1/2017\nN=37",
                                   "NHI introduced in \nQ3/2017\nN=24")),
         outcome2 = factor(outcome, 
                           levels = var_list, 
                           labels = var_label_l),
         pctchange = if_else(event.time != 0, NA, pctchange)) %>%
  group_by(group) %>%
  mutate(x.pctchange = max(event.time) - 0.8) %>% 
  filter(name == "uc.cls")

write_xlsx(plot.grp, paste0("./output/", "sfig_07_did_group.xlsx"))

# Plot -------------------------------------------------------------------------

xl <- range(plot.grp$event.time) # x-axis length / index

plot <- ggplot(
  plot.grp, aes(x = event.time, y = estimate)) + 
  facet_grid(outcome2 ~ group, scales = "free", switch = "y") +
  scale_x_continuous(name = "Time since NHI introduction, quarter",
                     breaks = xl[1]:xl[2]) +
  layer.64 +  # Add the layers
  theme.64()  # Add the theme

png(paste0("./graph/","sfig_07_did_group.png"), 
    units = "in", width = 4.75, height = 7, res = 500)
print(plot)
dev.off()

```
<br>

# Supplementary Figure 8. DiD analysis with adjusted model specifications 
# (balanced, last-treated, no covariate)
<br>
```{r sfig_8_adjust, eval=FALSE, include=FALSE}

# Load data
plot.dyn <- var_list %>%
  map(~ read_excel(paste0("./output/q_disl/", .x, "_aggte_dyn.xlsx"))) %>%
  bind_rows()

plot.dyn.b <- var_list %>%
  map(~ read_excel(paste0("./output/q_disl/", .x, "_aggte_dyn_b.xlsx"))) %>%
  bind_rows()

plot.dyn.b.new <- plot.dyn %>% # into new frame
  select(type, term, event.time, name, pretrendtype, outcome) %>%
  left_join(plot.dyn.b, by = c("type", "term", "event.time", 
                               "name", "pretrendtype", "outcome")) %>%
  filter(name == "uc.cls" | name == "vc.cls") %>% 
  mutate(name = ifelse(name == "uc.cls","uc.clsbal", "vc.clsbal"))

plot.ajt <- rbind(plot.dyn, plot.dyn.b.new) %>%
  filter(name %in% c("uc.cls", "uc.clsbal", "uc.clslast", "u.cls")) %>%
  mutate(outcome2 = factor(outcome, 
                           levels = var_list, 
                           labels = var_label_l),
         across(c(overall.att, overall.se, 
                  overall.conf.low, overall.conf.high), 
                ~ ifelse(event.time < 0, NA, .)),
         pctchange = scales::percent(overall.change, accuracy = 0.1),
         pctchange = if_else(event.time != 0, NA, pctchange),
         ajttype = factor(
           name,
           levels = c("uc.cls", "uc.clsbal", "uc.clslast", "u.cls"),
           labels = c("Base model:\nLast + not-yet-treated\nno anticipation",
                      "Balanced\nNHI exposure\n(e=0-2)",
                      "Last-treated units\nas control",
                      "No covariate\n(unconditional PTA)"))) %>%
  group_by(name) %>%
  mutate(x.pctchange = max(event.time) - 1.1)

write_xlsx(plot.ajt, paste0("./output/", "sfig_08_did_adjust.xlsx"))


# Plot -------------------------------------------------------------------------
 
xl <- range(plot.ajt$event.time) # x-axis length / index

plot <- ggplot(
  plot.ajt, aes(x = event.time, y = estimate)) + 
  facet_grid(outcome2 ~ ajttype, scales = "free", switch = "y") +
  scale_x_continuous(name = "Time since NHI introduction, quarter",
                     breaks = xl[1]:xl[2]) +
  layer.64 +  # Add the layers
  theme.64()  # Add the theme

png(paste0("./graph/", "sfig_08_did_adjust.png"), 
    units = "in", width = 4.75, height = 7, res = 500)
print(plot)
dev.off()

```
<br>

# Supplementary Figure 9. Difference-in-differences analysis using alternative heterogeneity-robust estimators
<br>
```{r sfig_9_hetero, eval=FALSE, include=FALSE}

data.htr <- read_excel("./output/sfig_09_did_hetero.xlsx")
unique(data.htr$method)

data.htr <- data.htr %>%
  mutate(outcome = factor(outcome, levels = var_list, labels = var_label_t))

# Plot -------------------------------------------------------------------------

xl <- range(data.htr$event.time) # x-axis length / index

plot <- ggplot(
  data.htr, aes(x = event.time, y = estimate, color = method)) + 
  facet_wrap(~ outcome, ncol = 2, scales = "free") +
  geom_vline(aes(xintercept = 0), linetype = "dashed", color = "black", 
             alpha = 0.9, linewidth = 0.05) +
  geom_hline(aes(yintercept = 0), color = "black", linewidth = 0.175) +
  geom_point(position = position_dodge2(width = 0.6), size = 0.6) +
  geom_linerange(aes(x = event.time, ymin = conf.low, ymax = conf.high), 
                 position = position_dodge2(width = 0.6), size = 0.3) +
  scale_y_continuous(name = "Difference-in-differences estimates") +
  scale_x_continuous(name = "Time since NHI introduction, quarter", 
                     breaks = xl[1]:xl[2]) +
  scale_color_manual(name = "Estimation Method", 
                     values = met.brewer("Cross", 6, "discrete")) +
  theme_bw() +
  theme(
    text = element_text(family = "Source Sans Pro"),
    panel.border = element_rect(color = "black", linewidth = 0.4),  
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(t = 0.2, r = 0.2, b = 0.1, l = 0.1, unit = "in"),
    axis.title.x = element_text(color = "black", size = 7, 
                                margin = margin(t = 0.15, unit = "in")),
    axis.title.y = element_text(color = "black", size = 7),
    axis.ticks = element_line(color = "black", linewidth = 0.2),
    axis.text = element_text(color = "black", size = 7),
    strip.text = element_text(color = "black", size = 7),
    strip.text.x = element_text(margin = margin(t = 0.15, b = 0.05, unit = "in")),
    strip.background = element_blank(),
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.text = element_text(size = 6),
    legend.key.size = unit(0.2, "in"),
    legend.key.spacing.y = unit(0.01, "in"),
    legend.margin = margin(l = -0.4, unit = "in")) +
  guides(col = guide_legend(ncol = 3, byrow = TRUE))

png(paste0("./graph/", "sfig_09_did_hetro.png"), 
    units = "in", width = 5.1, height = 6, res = 500)
print(plot)
dev.off()

```
<br>

# Supplementary Figure 10. DiD analysis with count data
<br>
```{r sfig_10_count, eval=FALSE, include=FALSE}

# Load data
var_list_c <- c("opo5", "opu5", "ipo5", "ipu5", "ipday", "del")
var_label_c <- c("Outpatient\nvisits,\naged ≥ 5",
                 "Inpatient\nadmissions,\naged ≥ 5",
                 "Outpatient\nvisits,\naged < 5",
                 "Inpatient\nadmissions,\naged < 5",
                 "Inpatient\ndays",
                 "Institutional\nbirth")

plot.dst <- var_list_c %>%
  map(~ read_excel(paste0("./output/q_disl/", .x, "_aggte_dyn.xlsx"))) %>%
  bind_rows() %>%
  mutate(facl = "dist")

read_data <- function(fl, vl) {
  map_dfr(vl, ~ {
    filepath <- paste0("./output/q_facl/", fl, "/", .x, "_aggte_dyn_", fl, ".xlsx")
    read_excel(filepath) %>%
      mutate(facl = fl)})
}

plot.cnt <- bind_rows(
  map_dfr(c("all", "dh", "hc"), ~ read_data(.x, var_list_c)),
  plot.dst)

plot.cnt <- plot.cnt %>%
  mutate(
    outcome2 = factor(outcome, 
                      levels = var_list_c, 
                      labels = var_label_c),
    facl2 = factor(facl, 
                   levels = c("all", "dist", "dh", "hc"),
                   labels = c("All (PH+DH+HC)", "District-level (DH+HC)", 
                              "District hospitals", "Health centers")),
    across(c(overall.att, overall.se, overall.conf.low, overall.conf.high), 
           ~ ifelse(event.time < 0, NA, .)),
    pctchange = scales::percent(overall.change, accuracy = 0.1),
    pctchange = if_else(event.time != 0, NA, pctchange),
    x.pctchange = max(event.time) - 1.1) %>%
  filter(name == "uc.cls")

write_xlsx(plot.cnt, paste0("./output/","sfig_10_did_count.xlsx"))

# Plot -------------------------------------------------------------------------

xl <- range(plot.cnt$event.time) # x-axis length / index

plot <- ggplot(
  plot.cnt, aes(x = event.time, y = estimate)) + 
  facet_grid(outcome2 ~ facl2, scales = "free", switch = "y") +
  scale_x_continuous(name = "Time since NHI introduction, quarter",
                     breaks = xl[1]:xl[2]) +
  layer.64 +  # Add the layers
  theme.64()  # Add the theme

png(paste0("./graph/", "sfig_10_did_count.png"), 
    units = "in", width = 4.75, height = 7, res = 500)
print(plot)
dev.off()

```
<br>

# Supplementary Figure 11. DiD analysis using monthly data
<br>
```{r sfig_11_month, eval=FALSE, include=FALSE}

# Load data
plot.dyn <- var_list %>%
  map(~ read_excel(paste0("./output/m_disl/", .x, "_aggte_dyn.xlsx"))) %>%
  bind_rows()

plot.dyn.b <- var_list %>%
  map(~ read_excel(paste0("./output/m_disl/", .x, "_aggte_dyn_b.xlsx"))) %>%
  bind_rows()

plot.dyn.b.new <- plot.dyn %>% # into new frame
  select(type, term, event.time, name, pretrendtype, outcome) %>%
  left_join(plot.dyn.b, by = c("type", "term", "event.time", 
                               "name", "pretrendtype", "outcome")) %>%
  filter(name == "uc.cls" | name == "vc.cls") %>% 
  mutate(name = ifelse(name == "uc.cls", "uc.clsbal", "vc.clsbal"))

plot.ajt <- rbind(plot.dyn, plot.dyn.b.new) %>%
  filter(name %in% c("uc.cls", "uc.clsbal", "uc.clslast", "u")) %>%
  mutate(outcome2 = factor(outcome, 
                           levels = var_list, 
                           labels = var_label_l),
         across(c(overall.att, overall.se, 
                  overall.conf.low, overall.conf.high), 
                ~ ifelse(event.time < 0, NA, .)),
         pctchange = scales::percent(overall.change, accuracy = 0.1),
         pctchange = if_else(event.time != 0, NA, pctchange),
         ajttype = factor(
           name,
           levels = c("uc.cls", "uc.clsbal", "uc.clslast", "u"),
           labels = c("Base model:\nLast + not-yet-treated\nno anticipation",
                      "Balanced\nNHI exposure\n(e=0-7)",
                      "Last treated units\nas control",
                      "No covariate\n(unconditional PTA)"))) %>%
  group_by(name) %>%
  mutate(x.pctchange = max(event.time) - 4) # higher number -> move to the left

write_xlsx(plot.ajt, paste0("./output/","sfig_11_did_month.xlsx"))


# Plot -------------------------------------------------------------------------
 
xl <- range(plot.ajt$event.time) # x-axis length / index

plot <- ggplot(
  plot.ajt, aes(x = event.time, y = estimate)) + 
  facet_grid(outcome2 ~ ajttype, scales = "free", switch = "y") +
  scale_x_continuous(name = "Time since NHI introduction, month",
                     breaks = c(-20, -10, 0, 10)) +
  layer.64 +  # Add the layers
  theme.64()  # Add the theme

png(paste0("./graph/", "sfig_11_did_month.png"), 
    units = "in", width = 4.75, height = 7, res = 500)
print(plot)
dev.off()

```

# Supplementary Figure 12. Robustness to violation of parallel trends
<br>
```{r sfig_12_honest, eval=FALSE, include=FALSE}

# Load data
var_list_h <- var_list[1:3] # "opo5r" "ipo5r" "opu5r"
var_label_h <- var_label_l[1:3]

plot.hon <- var_list_h %>%
  map(~ read_excel(paste0("./output/q_disl/", .x, "_honestdid.xlsx"))) %>%
  bind_rows() %>%
  mutate(
    outcome2 = factor(outcome, 
                      levels = var_list_h, 
                      labels = var_label_h),
    type2 = factor(type, levels = c("relative_magnitude", "smoothness"),
                   labels = c("Relative~magnitudes~(bar(M))",
                              "Smoothness~restrictions~(M)")),
    org = factor(ifelse(method == "Original", "Original", 
                        "Robust Confidence Interval"),
                 levels = c("Robust Confidence Interval", "Original")),
    M = ifelse(is.na(M) & type == "relative_magnitude", 0, M)) %>%
  dplyr::arrange(desc(type), outcome, name, desc(method), M) %>%
  mutate(
    seq = as.numeric(ave(ub, type, name, outcome, FUN = seq_along))) %>%
  dplyr::arrange(desc(type), outcome, name, desc(seq)) %>%
  group_by(type, name, outcome) %>%
  mutate(group_id = cur_group_id()) %>%
  ungroup()

smexp <- plot.hon %>% 
  dplyr::select(c("M", "group_id", "seq", "type")) %>%
  filter(type == "smoothness") %>%
  spread(., key = seq, value = M) %>%
  mutate(`1` = `2` - `3`) %>%
  gather(., seq, value,`1`:`11`) %>%
  mutate(seq = as.integer(seq))
colnames(smexp) <- c("group_id", "type", "seq", "M2")

plot.hon <- plot.hon %>% 
  left_join(smexp, by = c("group_id", "type", "seq")) %>%
  mutate(M2 = coalesce(M2, M)) %>% 
  filter(name == "uc.cls")

write_xlsx(plot.hon, paste0("./output/", "sfig_12_did_honest.xlsx"))

# Plot -------------------------------------------------------------------------

xl <- c(0, max(plot.hon$M2, na.rm=T)) # x-axis length / index

plot <- ggplot(
  plot.hon, aes(color = org)) + 
  ggh4x::facet_grid2(outcome2 ~ type2, scales = "free", axes = "x", 
                     independent = "x", switch = "both",
                     labeller = labeller(outcome2 = label_value,
                                         type2 = label_parsed)) +
  geom_hline(aes(yintercept = 0), color = "black", linewidth = 0.175) +
  geom_linerange(aes(x = M2, ymin = ub, ymax = lb), size = 0.3) +
  scale_x_continuous(name = "Restrictions") +   
  scale_color_manual(labels = c("Robust Confidence Interval", "Original"), 
                     values = c("#FC4E07", "#0CAFBB")) +
  theme_bw() +
  theme(
    text = element_text(family = "Source Sans Pro"),
    panel.border = element_rect(color = "black", linewidth = 0.4),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    plot.title = element_text(color = "black", size = 8),
    plot.margin = margin(t = 0.1, r = 0.1, b = 0.05, l = 0, unit = "in"),
    axis.title.x = element_text(color = "black", size = 6, 
                                margin = margin(0.15, 0, 0, 0, "in")),
    axis.title.y = element_blank(),
    axis.ticks.y = element_line(color = "black", linewidth = 0.2),
    axis.ticks.x = element_line(color = "black", linewidth = 0.2),
    axis.text.y = element_text(color = "black", size = 6),
    axis.text.x = element_text(color = "black", size = 6),
    strip.placement = "outside",
    strip.text = element_text(color = "black", size = 6),
    strip.text.y.left = element_text(angle = 0, hjust = 0, 
                                     margin = margin(t = 0, r = 0, 
                                                     b = 0, l = 0.05, "in")),
    strip.text.x = element_text(margin = margin(0.03, 0, 0.03, 0, "in")),
    strip.background = element_blank(),
    legend.title = element_blank(),
    legend.text = element_text(size = 4.75),
    legend.key.size = unit(0.15, "in"),
    legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "in"),
    legend.position = c(0.08, -0.19)) +
  guides(col = guide_legend(nrow = 1, reverse = TRUE)) # Reverse legend order

png(paste0("./graph/", "sfig_12_did_honest.png"), 
    units = "in", width = 4, height = 3.5, res = 500)
print(plot)
dev.off()

```
<br>

# Supplementary Figure 13. Controlled interrupted time series analyses
<br>
```{r sfig_13_cits, eval=FALSE, include=FALSE}

its_meta_final_q <- read_excel("./output/q_disl/cits_q.xlsx")
its_meta_final_m <- read_excel("./output/m_disl/cits_m.xlsx")

its_meta_final <- rbind(its_meta_final_q, its_meta_final_m)

plot.its <- its_meta_final %>%
  filter(!dataset %in% c("model_3_q", "model_3_m")) %>%
  filter(!(outcome %in% c("ipu5r", "opu5r", "iplos", "delr") & 
             dataset %in% c("model_2_q", "model_2_m"))) %>%
  mutate(outcome2 = factor(outcome, 
                           levels = var_list, 
                           labels = var_label_t),
         label = factor(label, 
                        levels = c("DID of slope (beta7)", 
                                   "DID of level (beta6)", 
                                   "Parallel trend (beta5)")),
         model = factor(dataset, 
                           levels = c("model_2_m", "model_1_m", 
                                      "model_2_q", "model_1_q"),
                           labels = c("Never-treated, monthly", 
                                      "Last-treated, monthly", 
                                      "Never-treated, quarterly", 
                                      "Last-treated, quarterly")))

write_xlsx(plot.its, "./output/sfig_13_cits.xlsx")

plot <- plot.its %>%
  ggplot(aes(y = label, x = estimate, color = model, shape = model)) + 
  geom_point(size = 0.8, position = position_dodge2(width = 0.5)) +
  geom_linerange(size = 0.3,aes(xmin = lowerci, xmax = upperci),
                 position = position_dodge2(width = 0.5)) +
  geom_vline(xintercept = 0, linetype = "dashed", size = 0.3) +
  facet_wrap(.~outcome2, ncol = 2, scales = "free") + #
  scale_y_discrete(
    labels = c("Parallel trend (beta5)" =
                 expression("Parallel trend ("*beta[5]*")"),
               "DID of level (beta6)" = 
                 expression("DID of level ("*beta[6]*")"),
               "DID of slope (beta7)" = 
                 expression("DID of slope ("*beta[7]*")"))) +
  scale_color_manual(
    name = "Control group, data frequency:",
    labels = c("Never-treated, monthly",
               "Last-treated, monthly",
               "Never-treated, quarterly",
               "Last-treated, quarterly"),
    values = c("#0CAFBB", "#0CAFBB", "#FC4E07", "#FC4E07")) +
  scale_shape_manual(
    name = "Control group, data frequency:",
    labels = c("Never-treated, monthly",
             "Last-treated, monthly",
             "Never-treated, quarterly",
             "Last-treated, quarterly"),
    values = c(15, 17, 15, 17)) +
  scale_x_continuous(name = "Difference-in-differences estimates") + 
  theme_bw() +
  theme(
    text = element_text(family = "Source Sans Pro"),
    panel.border = element_rect(color = "black", linewidth = 0.4),  
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    plot.title = element_text(color = "black",size=8),
    plot.margin = margin(t = 0.2, r = 0.2, b = 0.1, l = 0.1, unit = "in"),
    axis.title.x = element_text(color = "black",size = 7,
                                margin = margin(0.15, 0, 0, 0, "in")),
    axis.title.y = element_blank(),
    axis.ticks.y = element_line(color = "black", linewidth = 0.2),
    axis.ticks.x = element_line(color = "black", linewidth = 0.2),
    axis.text.y = element_text(color = 'black', size = 7),
    axis.text.x = element_text(color = 'black', size = 7),
    strip.text = element_text(color = 'black', size = 7),
    strip.text.x = element_text(margin = margin(0.15, 0, 0.05, 0, "in")),
    strip.background = element_blank(),
    legend.position = 'bottom',
              legend.title = element_text(size = 6),
              legend.text = element_text(size = 6),
              legend.key.size = unit(0.2, "in"),
              legend.key.spacing.y = unit(0.001, "in"),
              legend.margin = margin(t = 0, r = 0, b = 0, l = -0.1, 
                                     unit = "in")) +
  guides(col = guide_legend(ncol = 2, reverse = TRUE),
         shape = guide_legend(ncol = 2, reverse = TRUE))

png(paste0("./graph/", "sfig_13_cits.png"), 
      units = "in", width = 5.5, height = 6, res = 500)
print(plot)
dev.off()

```
